---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(dplyr)
library(lz4lite)
```

# lz4lite

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
<!-- badges: end -->

`lz4lite` provides access to the extremely fast compression in [lz4](https://github.com/lz4/lz4)
for performing in-memory compression.

The scope of this package is limited - it aims to provide functions for direct hashing of vectors
which contain raw, integer, real, complex or logical values.  It does this by operating on the
data payload within the vectors, and gains signficant speed by not serializing the R object
itselsf.  If you wanted to 
compress arbitrary R objects, you must first manually convert into a raw
vector representation using `base::serialize()`.

For a more general solution to fast serialization of R objects, see the 
[fst](https://github.com/fstpackage/fst) or [qs](https://cran.r-project.org/package=qs) packages.

Currently lz4 code provided with this package is v1.9.3.


### Design Choices

`lz4lite` will compress the *data payload* within a numeric-ish vector, and not the
R object itself.

### Limitations

* As it is the *data payload* of the vector that is being compressed, this
  does not include any notion of the container for that data i.e
  dimensions or other attributes are not compressed with the data.
* Values must be of type: raw, integer, real, complex or logical.
* Decompressed values are always returned as a vector i.e. all dimensional 
  information is lost during compression.


### What's in the box

* `lz4compress()` 
    - compress the data within a vector of  raw, integer, real, complex or logical values
    - set `use_hc = TRUE` to use the High Compression variant of LZ4. This
      variant can be slow to compress, but with higher compression ratios, 
      and it retains the fast decompression speed i.e. multiple gigabytes per second!
* `lz4decompress()` - decompress a compressed representation that was created with
  `lz4compress()`


### Installation

You can install from [GitHub](https://github.com/coolbutuseless/lz4lite) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/lz4lite)
```



## Compressing 1 million Integers

`lz4lite` supports the direct compression of raw, integer, real, complex and logical vectors.

On this test data, compression speed is ~600 MB/s, and decompression speed is ~3GB/s

```{r}
library(lz4lite)

N             <- 1e6
input_ints    <- (sample(seq(1:5), N, prob = (1:5)^2, replace = TRUE))
compressed_lo <- lz4_compress(input_ints)
compressed_hi <- lz4_compress(input_ints, use_hc = TRUE, hc_level = 12)
```


```{r echo = FALSE}
lens <- c(
  length(lz4_compress(input_ints, acc   =   1)),
  length(lz4_compress(input_ints, acc   =  10)),
  length(lz4_compress(input_ints, acc   =  20)),
  length(lz4_compress(input_ints, acc   =  40)),
  length(lz4_compress(input_ints, acc   = 100)),
  length(lz4_compress(input_ints, use_hc = TRUE, hc_level =   1)),
  length(lz4_compress(input_ints, use_hc = TRUE, hc_level =   2)),
  length(lz4_compress(input_ints, use_hc = TRUE, hc_level =   4)),
  length(lz4_compress(input_ints, use_hc = TRUE, hc_level =   8)),
  length(lz4_compress(input_ints, use_hc = TRUE, hc_level =  12))
)
```



<details>
<summary> Click here to show/hide benchmark code </summary>

```{r}
library(lz4lite)

res <- bench::mark(
  lz4_compress(input_ints, acc   =   1),
  lz4_compress(input_ints, acc   =  10),
  lz4_compress(input_ints, acc   =  20),
  lz4_compress(input_ints, acc   =  50),
  lz4_compress(input_ints, acc   = 100),
  lz4_compress(input_ints, use_hc = TRUE, hc_level =   1),
  lz4_compress(input_ints, use_hc = TRUE, hc_level =   2),
  lz4_compress(input_ints, use_hc = TRUE, hc_level =   4),
  lz4_compress(input_ints, use_hc = TRUE, hc_level =   8),
  lz4_compress(input_ints, use_hc = TRUE, hc_level =  12),
  check = FALSE
)
```
</details>


```{r echo = FALSE}
res %>% 
  mutate(`MB/s` = round(N*4/1024^2 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  mutate(compression_ratio = round(lens/(N*4), 3)) %>%
  select(expression, median, `itr/sec`, `MB/s`, compression_ratio) %>%
  knitr::kable()
```


### Decompressing 1 million integers

Decompression speed varies slightly depending upon the compressed size.

<details>
<summary> Click here to show/hide benchmark code </summary>

```{r}
res <- bench::mark(
  lz4_decompress(compressed_lo),
  lz4_decompress(compressed_hi)
)
```

</details>

```{r echo = FALSE}
res %>% 
  mutate(`MB/s` = round(N*4/1024^2 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  select(expression, median, `itr/sec`, `MB/s`)  %>%
  knitr::kable()
```



## Technical bits

### Why only vectors of raw, integer, real, complex or logical?

R objects can be considered to consist of: 

* a header - giving information like length and information for the garbage collector
* a body - data of some kind.

The vectors supported by `lz4lite` are those vectors whose body consists of 
data that is directly interpretable as a contiguous sequence of bytes representing 
numerical values.

Other R objects (like lists or character vectors) are really collections of pointers 
to other objects, and do not live in memory as a contiguous sequence of byte data.

### How it works.

##### Compression

1. Given a pointer to a standard numeric vector from R (i.e. an *SEXP* pointer).
2. Ignore any attributes or dimension information- just compress the data payload within 
   the object.
3. Prefix the compressed data with an 8 byte header giving size and SEXP type.
4. Return a raw vector to the user containing the compressed bytes.

##### Decompression

1. Strip off the 8-bytes of header information.
2. Feed the other bytes in to the LZ4 decompression function written in C
3. Use the header to decide what sort of R object this is.
4. Decompress the data into an R object of the correct type.
5. Return the R object to the user.

**Note:** matrices and arrays may also be passed to `lz4_compress()`, but since
no attributes are retained (e.g. dims), the uncompressed object
returned by `lz4_decompress()` can only be a simple vector.


### Framing of the compressed data

* `lz4lite` does **not** use the standard LZ4 frame to store data.
* The compressed representation is the compressed data prefixed with 
  a custom 8 byte header consisting of 
    * 'LZ4'
    * 1-byte for SEXP type i.e. INTSXP, RAWSXP, REALSXP or LGLSXP
    * 4-bytes representing an integer i.e. the number of bytes in the 
      original uncompressed data.
* This data representation 
    * is not compatible with the standard LZ4 frame 
      format.
    * is likely to evolve (so currently do not plan on compressing something in 
      one version of `lz4lite` and decompressing in another version.)








## Related Software

* [lz4](https://github.com/lz4/lz4) and [zstd](https://github.com/facebook/zstd) - both by Yann Collet
* [fst](https://github.com/fstpackage/fst) for serialisation of data.frames using
  lz4 and zstd
* [qs](https://cran.r-project.org/package=qs) for fast serialization of arbitrary R objects
  with lz4 and zstd

## Acknowledgements

* Yann Collett for releasing, maintaining and advancing 
[lz4](https://github.com/lz4/lz4) and [zstd](https://github.com/facebook/zstd)
* R Core for developing and maintaining such a wonderful language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
