---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(dplyr)
library(lz4lite)
```

# lz4lite

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

`lz4lite` provides access to the extremely fast compression in [lz4](https://github.com/lz4/lz4)
for performing in-memory compression.

As of v0.2.0, `lz4lite` can now serialize and compress any R object understood by
`base::serialize()`.

If the input is known to be an atomic, numeric vector, and you do not care about any attributes or
names on this vector, then `lz4_compress()`/`lz4_uncompress()` can be used. These
are bespoke serialization routines for atomic numeric vectors that run faster since
they avoid R's internals.

For a more general solution to fast serialization of R objects, see the 
[fst](https://github.com/fstpackage/fst) or [qs](https://cran.r-project.org/package=qs) packages.

Currently lz4 code provided with this package is v1.9.3.


### What's in the box

* **For arbitrary R objects**
    * `lz4_serialize`/`lz4_unserialize` serialize and compress any R object.
* **For atomic vectors with numeric values**
    * `lz4_compress()`/`lz4_uncompress()`
        - compress the data within a vector of  raw, integer, real, complex or logical values
        - faster than `lz4_serialize/unserialize` but throws away all attributes i.e. names, dims etc

### Installation

You can install from [GitHub](https://github.com/coolbutuseless/lz4lite) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/lz4lite)
```



## Basic usage of lz4lite

```{r}
dat <- mtcars


buf <- lz4_serialize(dat)
length(buf) # Number of bytes

# compression ratio
length(buf)/length(serialize(dat, NULL))

head(lz4_unserialize(buf))
```



## Compressing 1 million Integers


```{r}
library(lz4lite)

max_hc <- 12

set.seed(1)
N                <- 5e6
input_ints       <- sample(1:3, N, prob = (1:3)^3, replace = TRUE)
serialize_lo     <- lz4_serialize(input_ints, acceleration = 1)
serialize_hi_3   <- lz4hc_serialize(input_ints, level =  3)
serialize_hi_9   <- lz4hc_serialize(input_ints, level =  9)
serialize_hi_12  <- lz4hc_serialize(input_ints, level = max_hc)
compress_lo      <- lz4_compress(input_ints, acceleration = 1)
compress_hi_3    <- lz4hc_compress(input_ints, level = 3)
compress_hi_9    <- lz4hc_compress(input_ints, level = 9)
compress_hi_12   <- lz4hc_compress(input_ints, level = max_hc)
```


```{r echo = FALSE}
lens <- c(
  length(serialize(input_ints, NULL, xdr = FALSE)),
  length(lz4_serialize(input_ints, acceleration = 1)),
  length(lz4hc_serialize(input_ints, level =  3)),
  length(lz4hc_serialize(input_ints, level =  9)),
  length(lz4hc_serialize(input_ints, level = max_hc)),
  length(lz4_compress (input_ints, acceleration = 1)),
  length(lz4hc_compress (input_ints, level =  3)),
  length(lz4hc_compress (input_ints, level =  9)),
  length(lz4hc_compress (input_ints, level = max_hc))
)
```



<details>
<summary> Click here to show/hide benchmark code </summary>

```{r}
library(lz4lite)

res <- bench::mark(
  serialize(input_ints, NULL, xdr = FALSE),
  lz4_serialize(input_ints, acceleration = 1),
  lz4hc_serialize(input_ints, level =  3),
  lz4hc_serialize(input_ints, level =  9),
  lz4hc_serialize(input_ints, level = max_hc),
  lz4_compress (input_ints, acceleration = 1),
  lz4hc_compress (input_ints, level =  3),
  lz4hc_compress (input_ints, level =  9),
  lz4hc_compress (input_ints, level = max_hc),
  check = FALSE
)
```
</details>


```{r echo = FALSE}
res %>% 
  mutate(`MB/s` = round(N*4/1024^2 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  mutate(compression_ratio = round(lens/(N*4), 3)) %>%
  select(expression, median, `itr/sec`, `MB/s`, compression_ratio) %>%
  knitr::kable()
```


### uncompressing 1 million integers

uncompression speed varies slightly depending upon the compressed size.

<details>
<summary> Click here to show/hide benchmark code </summary>

```{r}
res <- bench::mark(
  lz4_uncompress(compress_lo),
  lz4_uncompress(compress_hi_3),
  lz4_uncompress(compress_hi_9),
  lz4_uncompress(compress_hi_12)
)
```

</details>

```{r echo = FALSE}
res %>% 
  mutate(`MB/s` = round(N*4/1024^2 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  select(expression, median, `itr/sec`, `MB/s`)  %>%
  knitr::kable()
```




### uncompressing 1 million integers

uncompression speed varies slightly depending upon the compressed size.

<details>
<summary> Click here to show/hide benchmark code </summary>

```{r}
res <- bench::mark(
  lz4_unserialize(serialize_lo),
  lz4_unserialize(serialize_hi_3),
  lz4_unserialize(serialize_hi_9),
  lz4_unserialize(serialize_hi_12)
)
```

</details>

```{r echo = FALSE}
res %>% 
  mutate(`MB/s` = round(N*4/1024^2 / as.numeric(median), 1)) %>%
  mutate(`itr/sec` = round(`itr/sec`)) %>%
  select(expression, median, `itr/sec`, `MB/s`)  %>%
  knitr::kable()
```






## Technical bits


### Framing of the compressed data

* `lz4lite` does **not** use the standard LZ4 frame to store data.
* The compressed representation is the compressed data prefixed with 
  a custom 8 byte header consisting of 
    * 3 bytes = 'LZ4' 
    * If this was produced with `lz4_serialize()` the next byte is 0x00, 
      otherwise it is a byte representing the SEXP of the encoded object.
    * 4-byte length value i.e. the number of bytes in the 
      original uncompressed data.
* This data representation 
    * is not compatible with the standard LZ4 frame 
      format.
    * is likely to evolve (so currently do not plan on compressing something in 
      one version of `lz4lite` and uncompressing in another version.)





## Related Software

* [lz4](https://github.com/lz4/lz4) and [zstd](https://github.com/facebook/zstd) - both by Yann Collet
* [fst](https://github.com/fstpackage/fst) for serialisation of data.frames using
  lz4 and zstd
* [qs](https://cran.r-project.org/package=qs) for fast serialization of arbitrary R objects
  with lz4 and zstd

## Acknowledgements

* Yann Collett for releasing, maintaining and advancing 
[lz4](https://github.com/lz4/lz4) and [zstd](https://github.com/facebook/zstd)
* R Core for developing and maintaining such a wonderful language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
